Bridging the Gap Between Visual and Analytical Machine Learning Testing

### 2023-10-05 report/icst24.tex
ref(report): rename esecfse24 -> icst24

\documentclass[acmsmall,screen,review,anonymous]{acmart}

%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmcopyright}
\copyrightyear{2018}
\acmYear{2018}
\acmDOI{XXXXXXX.XXXXXXX}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

\usepackage{xcolor}
\usepackage{listings}
\usepackage{caption}
\usepackage{subcaption}

@@

@@

\author{Anon}

\begin{abstract}
% what is the problem?
% why is this an important problem to solve?
% what did we do?
% what did we find?
Testing ML systems is a highly interactive process which demands
  a human-in-the-loop approach. In addition to writing tests for the
  code base, practitioners are required to analyse and interpret
  several visualisations using their domain expertise to validate if
  an ML system satisfies the required set of functional and
  non-functional properties. Visualisations are frequently used to
  qualitatively assess various parts of an ML pipeline. However,
  implicit knowledge gained from visualisations must be translated to
  explicit analytical tests that fail when there is change in any
  component of the ML pipeline. We conduct an empirical analysis of
  Jupyter notebooks to catalogue the state-of-the-art mappings between
  ML visualisations and assertions. We mine Github to collect Jupyter Notebooks that contain assertions written in Python. We develop a novel methodology to identify 1764 notebooks which contain an assertion below a visualisation. We manually analyse the 1.7K notebooks and identify 551 notebooks where at least one pair of visualisation and assertion are semantically related to one another. We further read and understand 90 notebooks in depth to identify 26 qualitative and their corresponding quantitative testing patterns. We showcase five such patterns in this paper and plan to make the entire catalogue publicly available in the near future.

\end{abstract}

@@

@@

Recent years show a shift from chasing high correctness of ML models to ensuring that they are safe for application in safety-critical domains such as healthcare, finance and autonomous driving. With the advent of Large Language Models, the power of AI to directly influence the daily lives of humans is growing at a radical pace. As reports of "hallucinated" facts, racist predicts and autonomous car crashes steadily increase, ensuring that ML systems function as they were intended to, takes precedence.

@@

From the 551 notebooks, we further perform an in-depth analysis of 90 notebooks. We identify 26 unique visualisation and assertion pairs. We discuss five such pairs in this paper and plan to release the full catalogue publicly in the near future.

\end{description}

@@

\citeauthor{psallidas2019data} provide an overview of the evolving landscape of computational notebooks by analysing six million Python notebooks, two million enterprise data science pipelines, and source code and metadata from over 900 releases of 12 important data science libraries. Their findings can be used by system builders for improving data science tools and also by data science practitioners to understand the current trends and what technologies they should focus on. \citeauthor{pimentel2019large} mined 1.4 million notebooks from Github to conduct an empirical study on the good and bad coding practices in computational notebooks. Based on their analysis, the authors propose guidelines on improving reproducibility of computational notebooks. To enable future research on computational notebooks, \citeauthor{quaranta2021kgtorrent} mine Kaggle~\footnote{https://kaggle.com} and present \textit{KGTorrent}, a public dataset consisting of approximately 2.5 million Jupyter notebooks written in Python. The dataset also contains a relational database dump of metadata regarding publicly available notebooks on Kaggle.

@@

To manage and prune the multiple versions of code that accumulate over time when developing in notebooks, \citeauthor{head2019managing} propose a code gathering tool that allows practitioners to review and only keep the relevant version of code. Other tools such as \textit{WrangleDoc} and \textit{VizSmith} have also been proposed to aid ML practitioners when working within computational notebooks~\cite{yang2021subtle, bavishi2021vizsmith}.

@@

Prior studies have been conducted to understand how visual analytics techniques are currently being used in ML. \citeauthor{yuan2021survey} conduct a systematic review of 259 papers and propose a taxonomy of visual analytics techniques for ML. \citeauthor{hohman2019visual} conduct a survey of visual analytics techniques for Deep Learning Models. The findings of the study indicate that visual analytics has been widely adopted in ML for model explanation, interpretation, debugging and improvement.

Several tools have been proposed by the visual analytics research community to aid practitioners in understanding how their ML models operate. \citeauthor{wexler2020if} propose \textit{What-If}, a visual analytics tool to explore alternative hypotheses, generate counterfactual reasoning, investigate decision boundaries of the model and how change in data affects the model predictions. To reduce the cognitive load of ML practitioners during model building phase, \citeauthor{amershi2015modeltracker} propose \textit{ModelTracker}. Given a trained model and a sample dataset, ModelTracker presents all the information in traditional summary statistics along with the performance of the model.

@@

\begin{minipage}{0.45\textwidth}
  \begin{lstlisting}
{
  "metadata": {
    "kernel_info": {...},
    "language_info": {...},
  },
  "nbformat": 4,
  "nbformat_minor": 0,
  "cells": [ # dictionary of all cells in notebook ]
}
  \end{lstlisting}
  \captionof{lstlisting}{Top-level JSON format of Jupyter notebooks. See Listing~\ref{lst:cell-md}, \ref{lst:cell-code} and \ref{lst:cell-output} for internal structure of various cells.}
  \label{lst:nbformat}
\end{minipage}
\hfill
\begin{minipage}{0.45\textwidth}
  \begin{lstlisting}
{
  "cell_type": "markdown",
  "metadata": {},
  "source": [
    "# Hello world example\n",
    "The following cell contains python code for greeting people!\n"
  ]
}
  \end{lstlisting}
  \captionof{lstlisting}{JSON structure of markdown cells.}
  \label{lst:cell-md}
\end{minipage}
\begin{minipage}{0.45\textwidth}
  \begin{lstlisting}
{
  "cell_type": "code",
  "metadata": {},
  "source": [
    "def greet(name):\n",
    "    print(f'Hello, {name}!')\n"
  ]
}
  \end{lstlisting}
  \captionof{lstlisting}{JSON structure of code cells.}
  \label{lst:cell-code}
\end{minipage}
\hfill
\begin{minipage}{0.45\textwidth}
  \begin{lstlisting}
{
  "cell_type": "code",
  "metadata": {},
  "source": [...],
  "outputs": [
    {
      "output_type": "display_data",
      "data": {
        "image/png": "<base64-encoded-png-data>"
      }
    }
  ]
}
  \end{lstlisting}
  \captionof{lstlisting}{JSON structure of code cells with an image output.}
  \label{lst:cell-output}
\end{minipage}

@@

\begin{table}
\centering
\caption{Frequent examples of post-condition checks observed in this study.}
\begin{tabular}{p{0.3\textwidth} p{0.4\textwidth} p{0.3\textwidth}}
\toprule
\emph{\textbf{Visualisation}}&
\emph{\textbf{Post-condition check}}&
\emph{\textbf{Rationale}}\\
\midrule
\includegraphics[width=\linewidth]{post-cond-01.png}&
\begin{lstlisting}[language=Python]
assert f1.gca().has_data()
\end{lstlisting}&
The assertion checks if the output of the visualisation code cell contains an image.\\
\includegraphics[width=\linewidth]{post-cond-02.png}&
\begin{lstlisting}[language=Python]
assert_almost_equal(sigmoid(-50), 0, delta=1e-10)
assert_almost_equal(sigmoid(0), 0.5, delta=1e-10)
assert_almost_equal(sigmoid(50), 1, delta=1e-10)
\end{lstlisting}&
The assertions are spot-checking the sigmoid activation function.\\
\includegraphics[width=\linewidth]{post-cond-03.png}&
\begin{lstlisting}[language=Python]
# The axis should contain 5 bars
assert len(ax.containers[0]) == 5

# One of the x tick labels should be "blue"
tick_text = [tick.get_text() for tick in ax.get_xticklabels()]
assert "blue" in tick_text
\end{lstlisting}&
The asserts are validating the correctness of the visualisation itself.\\
\bottomrule
\end{tabular}
\label{tab:post-cond}
\end{table}

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{method.pdf}
  \caption{Methodology used to collect and analyse Jupyter notebooks
    in this study.}
  \label{fig:method}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{nb-structure.pdf}
  \caption{Arrangement of code cells containing assertion and
    visualisation.}
  \label{fig:cell-arrangement}
\end{figure}

@@

\begin{table}
  \centering
  \caption{Observations from random samples of notebooks with various
  cell arrangement.}
  \begin{tabular}{l p{0.4\textwidth} p{0.4\textwidth}}
    \toprule
    \emph{\textbf{ID}}&
    \emph{\textbf{Cell arrangement}} &
    \emph{\textbf{Observations}}\\
    \midrule
    1 &
    Assertion cell above visualisation cell &
    Least likely to be related to the visualisation below.\\
    \midrule
    \multicolumn{3}{l}{\textbf{Assertion and visualisation in the same code cell}}\\
    \midrule
    2.1 &
    Assertion above visualisation code &
    More likely to be a pre-condition check.\\
    2.2 &
    Assertion below visualisation code &
    More likely to be a post-condition check.\\
    \midrule
    \multicolumn{3}{l}{\textbf{Assertion cell below visualisation cell}}\\
    \midrule
    3.1 &
    Assertion cell more than 10 lines &
    Better signal-to-noise ratio compared to Arrangement 2.1. However, More likely to be post-condition check or unrelated to the visualisation above.\\
    3.2 &
    Assertion cell has max 10 lines &
    Most likely to be related to the visualisation above.\\
    \bottomrule
  \end{tabular}
  \label{tab:cell-arrangement}
\end{table}

@@

\begin{table}
  \centering
  \caption{Exclusion criteria used to identify related visualisation and assertion pairs.}
  \begin{tabular}{l l r}
    \toprule
    \textbf{ID} &
    \textbf{Exclusion Criteria} &
    \textbf{Num. excluded NBs}\\
    \midrule
    \textbf{EC1} &
    The visualisation and assertion are not related &
    440\\
    \textbf{EC2} &
    Notebook does not contain a trained ML model &
    383\\
    \textbf{EC3} &
    The assertion is a post-condition check &
    200\\
    \textbf{EC4} &
    Notebook not written in English &
    159\\
    \textbf{EC5} &
    Notebook contains incorrect code &
    31\\
    \midrule
    \multicolumn{2}{c}{\emph{\textbf{Total number of NBs excluded:}}} &
    1.213\\
    \bottomrule
  \end{tabular}
  \label{tab:exclusion-criteria}
\end{table}

@@

For the 551 notebooks where

We qualitatively assess the comments made for the
We qualitatively assess the comments made for all notebooks during the analysis conducted in phase two of the study, as described in Section~\ref{sec:identify-related-pairs}.

@@

% NOTE comment on the over-whelming presence of magic thresholds in ML testing! where are authors getting these numbers from? link this to requirements engineering?

@@

\bibliographystyle{ACM-Reference-Format}
\bibliography{bibliography}
\end{document}

>>>>

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{array}

@@

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

@@

\author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
\and
\IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
}

@@

\begin{abstract}
Testing ML systems is a highly interactive process which demands a
human-in-the-loop approach. In addition to writing tests for the code
base, practitioners are required to analyse and interpret several
visualisations using their domain expertise to validate if an ML
system satisfies the required set of functional and non-functional
properties. Visualisations are frequently used to qualitatively assess
various parts of an ML pipeline. However, implicit knowledge gained
from visualisations must be translated to explicit analytical tests
that fail when there is change in any component of the ML pipeline. We
conduct an empirical analysis of Jupyter notebooks to catalogue the
state-of-the-art mappings between ML visualisations and assertions. We
mine Github to collect Jupyter Notebooks that contain assertions
written in Python. We develop a novel methodology to identify 1764
notebooks which contain an assertion below a visualisation. We
manually analyse the 1.7K notebooks and identify 551 notebooks where
at least one pair of visualisation and assertion are semantically
related to one another. We further read and understand 90 notebooks in
depth to identify 26 qualitative and their corresponding quantitative
testing patterns. We showcase five such patterns in this paper and
plan to make the entire catalogue publicly available in the near
future.
\end{abstract}

\begin{IEEEkeywords}
  ML Testing, Visualisations, ???
\end{IEEEkeywords}

@@

Recent years show a shift from chasing high correctness of ML models
to ensuring that they are safe for application in safety-critical
domains such as healthcare, finance and autonomous driving. With the
advent of Large Language Models, the power of AI to directly influence
the daily lives of humans is growing at a radical pace. As reports of
``hallucinated'' facts, racist predicts and autonomous car crashes
steadily increase, ensuring that ML systems function as they were
intended to, takes precedence.

@@

From the 551 notebooks, we further perform an in-depth analysis of 90
    notebooks. We identify 26 unique visualisation and assertion
    pairs. We discuss five such pairs in this paper and plan to
    release the full catalogue publicly in the near future.
\end{description}

@@

\cite{psallidas2019data} provide an overview of the evolving landscape
of computational notebooks by analysing six million Python notebooks,
two million enterprise data science pipelines, and source code and
metadata from over 900 releases of 12 important data science
libraries. Their findings can be used by system builders for improving
data science tools and also by data science practitioners to
understand the current trends and what technologies they should focus
on. \cite{pimentel2019large} mined 1.4 million notebooks from Github
to conduct an empirical study on the good and bad coding practices in
computational notebooks. Based on their analysis, the authors propose
guidelines on improving reproducibility of computational notebooks. To
enable future research on computational notebooks,
\cite{quaranta2021kgtorrent} mine Kaggle~\footnote{https://kaggle.com}
and present \textit{KGTorrent}, a public dataset consisting of
approximately 2.5 million Jupyter notebooks written in Python. The
dataset also contains a relational database dump of metadata regarding
publicly available notebooks on Kaggle.

@@

To manage and prune the multiple versions of code that accumulate over
time when developing in notebooks, \cite{head2019managing} propose
a code gathering tool that allows practitioners to review and only
keep the relevant version of code. Other tools such as
\textit{WrangleDoc} and \textit{VizSmith} have also been proposed to
aid ML practitioners when working within computational
notebooks~\cite{yang2021subtle, bavishi2021vizsmith}.

@@

Prior studies have been conducted to understand how visual analytics
techniques are currently being used in ML. \cite{yuan2021survey}
conduct a systematic review of 259 papers and propose a taxonomy of
visual analytics techniques for ML. \cite{hohman2019visual} conduct
a survey of visual analytics techniques for Deep Learning Models. The
findings of the study indicate that visual analytics has been widely
adopted in ML for model explanation, interpretation, debugging and
improvement.

Several tools have been proposed by the visual analytics research
community to aid practitioners in understanding how their ML models
operate. \cite{wexler2020if} propose \textit{What-If}, a visual
analytics tool to explore alternative hypotheses, generate
counterfactual reasoning, investigate decision boundaries of the model
and how change in data affects the model predictions. To reduce the
cognitive load of ML practitioners during model building phase,
\cite{amershi2015modeltracker} propose \textit{ModelTracker}. Given
a trained model and a sample dataset, ModelTracker presents all the
information in traditional summary statistics along with the
performance of the model.

@@

\begin{minipage}{0.45\textwidth}
  \begin{lstlisting}
{
  "metadata": {
    "kernel_info": {...},
    "language_info": {...},
  },
  "nbformat": 4,
  "nbformat_minor": 0,
  "cells": [ # dictionary of all cells in notebook ]
}
  \end{lstlisting}
  \captionof{lstlisting}{Top-level JSON format of Jupyter notebooks.
    See Listing~\ref{lst:cell-md},~\ref{lst:cell-code}
    and~\ref{lst:cell-output} for internal structure of various
    cells.}\label{lst:nbformat}
\end{minipage}
\hfill
\begin{minipage}{0.45\textwidth}
  \begin{lstlisting}
{
  "cell_type": "markdown",
  "metadata": {},
  "source": [
    "# Hello world example\n",
    "The following cell contains python code for greeting people!\n"
  ]
}
  \end{lstlisting}
  \captionof{lstlisting}{JSON structure of markdown cells.}
  \label{lst:cell-md}
\end{minipage}
\begin{minipage}{0.45\textwidth}
  \begin{lstlisting}
{
  "cell_type": "code",
  "metadata": {},
  "source": [
    "def greet(name):\n",
    "    print(f'Hello, {name}!')\n"
  ]
}
  \end{lstlisting}
  \captionof{lstlisting}{JSON structure of code cells.}
  \label{lst:cell-code}
\end{minipage}
\hfill
\begin{minipage}{0.45\textwidth}
  \begin{lstlisting}
{
  "cell_type": "code",
  "metadata": {},
  "source": [...],
  "outputs": [
    {
      "output_type": "display_data",
      "data": {
        "image/png": "<base64-encoded-png-data>"
      }
    }
  ]
}
  \end{lstlisting}
  \captionof{lstlisting}{JSON structure of code cells with an image output.}
  \label{lst:cell-output}
\end{minipage}

@@

\begin{table*}
\centering
\caption{Frequent examples of post-condition checks observed in this study.}
  \begin{tabular}{m{0.3\textwidth} m{0.4\textwidth} m{0.3\textwidth}}
\hline
\emph{\textbf{Visualisation}}&
\emph{\textbf{Post-condition check}}&
\emph{\textbf{Rationale}}\\
\hline
\includegraphics[width=\linewidth]{post-cond-01.png}&
\begin{lstlisting}[language=Python]
assert f1.gca().has_data()
\end{lstlisting}&
The assertion checks if the output of the visualisation code cell contains an image.\\
\includegraphics[width=\linewidth]{post-cond-02.png}&
\begin{lstlisting}[language=Python]
assert_almost_equal(sigmoid(-50), 0, delta=1e-10)
assert_almost_equal(sigmoid(0), 0.5, delta=1e-10)
assert_almost_equal(sigmoid(50), 1, delta=1e-10)
\end{lstlisting}&
The assertions are spot-checking the sigmoid activation function.\\
\includegraphics[width=\linewidth]{post-cond-03.png}&
\begin{lstlisting}[language=Python]
# The axis should contain 5 bars
assert len(ax.containers[0]) == 5

# One of the x tick labels should be "blue"
tick_text = [tick.get_text() for tick in ax.get_xticklabels()]
assert "blue" in tick_text
\end{lstlisting}&
The asserts are validating the correctness of the visualisation itself.\\
\hline
\end{tabular}
\label{tab:post-cond}
\end{table*}

@@

\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{method.pdf}
  \caption{Methodology used to collect and analyse Jupyter notebooks
    in this study.}
  \label{fig:method}
\end{figure*}

@@

\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{nb-structure.pdf}
  \caption{Arrangement of code cells containing assertion and
    visualisation.}
  \label{fig:cell-arrangement}
\end{figure*}

@@

\begin{table}
  \centering
  \caption{Observations from random samples of notebooks with various
  cell arrangement.}
  \begin{tabular}{l p{0.2\textwidth} p{0.2\textwidth}}
    \hline
    \emph{\textbf{ID}}&
    \emph{\textbf{Cell arrangement}} &
    \emph{\textbf{Observations}}\\
    \hline
    1 &
    Assertion cell above visualisation cell &
    Least likely to be related to the visualisation below.\\
    \hline
    \multicolumn{3}{l}{\textbf{Assertion and visualisation in the same code cell}}\\
    \hline
    2.1 &
    Assertion above visualisation code &
    More likely to be a pre-condition check.\\
    2.2 &
    Assertion below visualisation code &
    More likely to be a post-condition check.\\
    \hline
    \multicolumn{3}{l}{\textbf{Assertion cell below visualisation cell}}\\
    \hline
    3.1 &
    Assertion cell more than 10 lines &
    Better signal-to-noise ratio compared to Arrangement 2.1. However, More likely to be post-condition check or unrelated to the visualisation above.\\
    3.2 &
    Assertion cell has max 10 lines &
    Most likely to be related to the visualisation above.\\
    \hline
  \end{tabular}
  \label{tab:cell-arrangement}
\end{table}

@@

\begin{table}
  \centering
  \caption{Exclusion criteria used to identify related visualisation and assertion pairs.}
  \begin{tabular}{l p{0.2\textwidth} r}
    \hline
    \textbf{ID} &
    \textbf{Exclusion Criteria} &
    \textbf{Num. excluded NBs}\\
    \hline
    \textbf{EC1} &
    The visualisation and assertion are not related &
    440\\
    \textbf{EC2} &
    Notebook does not contain a trained ML model &
    383\\
    \textbf{EC3} &
    The assertion is a post-condition check &
    200\\
    \textbf{EC4} &
    Notebook not written in English &
    159\\
    \textbf{EC5} &
    Notebook contains incorrect code &
    31\\
    \hline
    \multicolumn{2}{c}{\emph{\textbf{Total number of NBs excluded:}}} &
    1.213\\
    \hline
  \end{tabular}
  \label{tab:exclusion-criteria}
\end{table}

@@

We qualitatively assess the comments made for all notebooks during the
analysis conducted in phase two of the study, as described in
Section~\ref{sec:identify-related-pairs}.

@@

% NOTE comment on the over-whelming presence of magic thresholds in ML
% testing! where are authors getting these numbers from? link this to
% requirements engineering?

@@

\bibliographystyle{IEEEtran}
\bibliography{bibliography}

\end{document}
\endinput
